### 2012-06-29 app-gravity.tex

\section{Gravity solver}

\red{This needs to be fleshed out a lot - basically we want to emphasize how we solve the
root grid (mention parallelization strategy for FFT), how the potentials calculated are 
interpolated to subgrids for the multigrid solution on the subgrids, and how particles are 
actually moved.
}

\red{**********  Which version of the gravity solver will we be using for the public release?  
The new Dan Reynolds mulgrid-everywhere solver, or the one that's currently in the public
release?  Alter paper and other documentation accordingly.}

With and without cosmology - how does this affect what is solved/how it's solved?
Why do we do $\rho-\bar{\rho}$ instead of just $\rho$?  What is the finite difference
equation that you actually solve in the end?

How are particles interpolated to grids, and accelerations from grid
to particles?

\subsection{multigrid}
This is for James.

\subsection{leapfrog integrator}
Talk about how particles' accelerations are calculated, etc.