### 2012-06-29 app-ppmhydro.tex

\section{The PPM hydro method}\label{app.ppm}
\red{
\begin{enumerate}
  \item Cite \citet{1995CoPhC..89..149B} for Lagrangian Remap version
  \item State that Lagrangian Remap it doesn't really work and nobody uses it (why is that?)
  \item Brief overview of PPM: 
  \begin{enumerate}
    \item Parabolic reconstruction
    \item Order Lowering near shocks
    \item Characteristic tracing for Left and Right states
    \item I don't remember if we use a contact steepener:  If yes,
    describe.  If no, say so.
  \end{enumerate}
  \item Repeat bits of \citet{1984JCoPh..54..174C} that have been altered
        for cosmology 
  \item I think C\&W 1984 discuss gravity: double check we do the
  same.  If not, comment.
  \item Discus TwoShock solver, 
    \begin{enumerate} 
       \item Short method review, I think Woodward 1986.
       \item Discuss Rarefaction updates 
       \item Discuss alterations for Expansion
       \item Discuss alterations for Gravity (if not in Woodward 1986)
    \end{enumerate}
\end{enumerate}
}%end red