### 2012-06-29 discussion.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\red{
\begin{enumerate}
\item What physics will be put into the code in the next couple of years? (ideal mhd, FLD, ray tracing, more...)
\item What will never be in the code (non-cartesian coordinates, SPH)
\item What are clear areas of improvement?  (gravity solver, maybe more advanced/different hydro methods, parallelization, load balancing)
\item Discuss relative ease of adding new particle types (dust, ``galaxy particles,'' etc.) and some other new packages (cooling).  Other things are harder (MHD, radiation transport).  Outline what's easy and what's hard, in very general terms.
\end{enumerate}
} %end red