### 2013-01-18 development.tex
added two new sections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code Development Methodology}

\begin{enumerate}
\item evolution of development strategy
\item ideas for where it might go in the future
\item aim for no more than 1 page.
\end{enumerate}