### 2013-01-18 conclusions.tex
adding a bunch of new sections and subsections.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

\begin{enumerate}
\item near-future development plans (OpenMP, ...)
\item far-future development plans (enzo 3.0 and beyond)
\item clear areas of improvement
\end{enumerate}