### 2013-01-18 development.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code development methodology}

\begin{enumerate}
\item evolution of development strategy over time
\item where we might want that to go in the future
\item keep to no more than one ApJ page
\end{enumerate}