### 2013-01-18 numerical-radiation-fld.tex

\subsection{Radiation transport: Flux-limited diffusion}
\label{sec.num.rad-fld}