### 2013-01-18 numerical-radiation-raytracing.tex

\subsection{Radiation transport: ray tracing}
\label{sec.num.raytracing}