### 2013-01-22 code-tests-fld.tex
Added separate file for FLD test problem

\subsubsection{Representative FLD test}
\label{sec.tests.fld}

\red{(Dan)}
One test for the implicit FLD: best/hardest one.