### 2013-01-23 code-tests-fld.tex
Updated FLD section of method paper (section 3.13)

\subsubsection{Representative FLD test}
\label{sec.tests.fld}

\red{(Dan)}
One test for the implicit FLD: best/hardest one.

%% With regards to the test problem, we were hoping you could write
%% section 4.2.14, going along with the general structure of the test
%% problem explained at the beginning of section 4.2.  I don't anticipate
%% it being more than a couple of 1D or 2D simulations or a test problem
%% of your choice, a plot or two, and a couple of references to the
%% original Sedov/Taylor papers and your method paper.   If you could
%% also provide (by emailing me) the enzo parameter files you use for the
%% simulations that appear in the paper, as well as the scripts you use
%% to generate the figures shown in the paper, that would be great -
%% we're going to make a "method paper" option in the test suite so that
%% people can easily reproduce the figures if they want.