### 2013-03-15 numerical-starformation.tex
intermediate checkin.

\subsection{Star formation and feedback}
\label{sec.ov.star}

\subsubsection{Overview}
% Overview of methodology

While the physics discussed in other sections of this paper is crucial
to the study of cosmological structure formation, most observations of
galaxies in the infrared, optical, and ultraviolet wavelengths are of
stellar populations.  Furthermore, stars eject energy and metal-
enriched gas throughout their lives, drastically modifying their own
environments' cooling properties.  Broadly speaking, the formation of
galaxies and galaxy clusters cannot be completely modeled without
including the feedback of energy and metals from stellar populations.
In particular, it is thought that feedback is crucial for suppressing
the large numbers of dwarf-like galaxies that CDM theories predict
\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early burst of
star formation could remove a large fraction of cold gas from such
systems \citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the
unexpectedly low luminosities of small clusters and groups (relative
to the richest clusters) is commonly explained through feedback
\citep{1991ApJ...383..104K}.  Energy ejected during the formation of
the cluster ellipticals lowers the central density and hence the X-ray
luminosity of such clusters \citep{1997ApJ...484L..21C}.  For all of
these reasons, it is critical to include star formation and the
feedback of energy and metals into the interstellar and intergalactic
media in cosmological simulations.

Modeling star formation in the context of galaxy formation forces one
to confront two fundamental problems. The dynamical range required to
study galaxy formation and evolution is substantial, and attempting to
directly resolve the formation of individual stars in the context of a
Milky Way-type galaxy is computationally unfeasible.  Also, a strong
theoretical understanding of the details of star formation is
currently lacking.  Due to these challenges, models of star formation
in cosmological simulations are largely phenomenological, and the
``stars'' that are formed are in most cases meant to represent an
ensemble stellar population (although not all; see
Section~\ref{sec:starform_pop3}).
% BWO: I use 'phenomenological' right above this, but could probably
% go for 'parametric' or 'heuristic' if people object to my usage.

The \enzo\ code includes approximately ten models for star formation
and feedback.  Broadly speaking, these methods all work in similar
ways: At a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but may include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with the
same mass is placed in the center of that cell, with the same momentum
vector as the removed gas.  This star particle is then allowed to
inject mass, momentum, thermal energy, metals, and possibly magnetic
fields and/or cosmic ray populations into its local environment.  In
general, the particle is treated as an ensemble of stars, with
feedback properties occuring over time according to the assumed
initial mass function of the stellar population.

In the following sections, we describe some of the more widely-used star formation
and feedback methods used in \enzo.  We note that similar methods have
been employed in many other codes used for galaxy formation, with
comparable implementations for both star formation and feedback in
other grid-based codes.  With regards to particle-based codes, star
formation is broadly similar in implementation, though feedback is
typically implemented in a very different way due to the Lagrangian
nature of the method \citep[see, e.g.,][]{sh03a,sh03b,hs03}.

% Specific example: Cen & Ostriker
\subsubsection{Cen \& Ostriker}
\label{sec:starform_cen}

The Cen \& Ostriker method is a heuristic model of star formation
on galactic scales.  This method, first described by
\citet{CO1992}, assumes that star particles

s similar in some ways to the Kravtsov algorithm
but has more complex criteria for determining where stars should
be formed.  In this method, cells that form stars must have a 
baryon overdensity higher than some threshold 
$\rho_b/\bar{\rho}_b \geq \eta$ 
where $\rho_b$ is the baryon density in that cell, 
$\bar{\rho_b}$ is the mean baryon density in the simulation,
and $\eta$ is the user-defined overdensity threshold.
Additionally, the gas in the cells must be contracting, 
cooling rapidly, and gravitationally unstable, e.g.:

\begin{equation}
\nabla \cdot \vec{v}_b < 0,
\label{cencont}
\end{equation}

\begin{equation}
t_{cool} < t_{dyn} \equiv \sqrt{3 \pi / 32G \rho_{tot}},
\end{equation}

\begin{equation}
m_{b} > m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}C^{3}
\left[ 1 + \frac{\delta\rho_{d}}{\delta\rho_{b}} \right]^{-3/2}
\end{equation}

where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
$\rho_{d}$ are the cell's baryon and dark matter density, respectively,
$\rho_{total} = \rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the 
baryonic mass in the cell and jeans mass of the cell, and C is the 
isothermal sound speed in the cell.  
If all of these criteria are met, 
the mass of a star particle is calculated as \(m_{*} = m_{b} 
\frac{ \Delta t}{ t_{dyn} } f_{*eff} \), 
where $f_{*eff}$ is the star formation efficiency parameter.

If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle 
is created and given several attributes:  Mass, a unique index number, 
the time of formation $t_{form}$, the local dynamical free-fall time 
$t_{dyn}$ and the metallicity fraction of the baryon gas in the cell 
$f_{Zb}$.  There is a user-defined minimum dynamical time
$T_{dyn,min}$ which is observationally motivated and affects the
feedback rates (see below).
The particle is placed in the center of the cell and given 
the same peculiar velocity as the gas in the cell, and is then treated 
in the same manner as the dark matter particles.  An 
amount of baryon gas corresponding to the new particle's mass is 
removed from the cell.

In addition, we have added a stochastic star formation algorithm that keeps 
track of all of the sub-mass threshold stars that should have been created 
and when the total mass of uncreated stars is greater than the minimum mass, 
a star particle is created.

The feedback of energy and metals into the baryon gas is similar to
the Kravtsov feedback, with some important differences.  The star formation 
algorithm creates each star particle instantaneously.  However, feedback
should take place over a significant timescale, as all of the stars contained
within the ``star particle'' would form over a long period of time.  Therefore,
we assume that for the purposes of feedback that the mass of stars formed
at a time $t$ with finite timestep $\Delta t$ is:

\begin{eqnarray}  
\Delta m_{sf} =  m_{*} \left[ \left(1 + \frac{t-t_{form}}{t_{dyn}}  \right)
%\nonumber \times \\ % uncomment for 2-column
		   \exp{\left( \frac{-(t-t_{form})}{t_{dyn}} \right) }
%                    \nonumber \\ 
	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) 
%	       \nonumber \\ 
                    \exp{ \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right)  }
                   \right]
\end{eqnarray}

which can be represented more clearly in integral form:

\begin{eqnarray}
\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) 
%\nonumber \times \\ 
\exp{ \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) }
\end{eqnarray}

During this timestep, we assume that the star particle feeds back
metal-enriched gas and thermal energy from supernovae and from stellar winds.
Since massive stars have very short lifetimes, we assume that there is an
immediate feedback of some fraction $f_{SN}$ of the rest energy from the
stars created that timestep into the baryon gas, such that 
$E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of light.
In addition, a fraction $f_{Z*}$ of the metal from the star particle
is fed back into the baryon gas, which takes into account the effects
of metal recycling.  Finally, a fraction of the mass $f_{m*}$ from all
stars (rather than just supernovae) is fed back into the gas along with
momentum in order to simulate the mass ejection from non-exploding stars
via stellar winds.

There are six user-defined parameters in this algorithm:  three deal with the
star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and three 
deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of these
parameters are completely free, while others can be guided by observation
or theory.  For example, the supernova feedback parameter, $f_{SN}$, can be 
constrained assuming that, for every $200 M_\odot$ of stars created, one
supernova occurs, and this event feeds back approximately $10^{51}$ ergs of
thermal energy, giving:

\begin{equation}
f_{SN} = \frac{10^{51} ergs}{200~M_\odot c^2} \simeq 3 \times 10^{-6}
\end{equation}

The metal yield $f_{Z*}$, defined as the mass in metals produced per unit
mass of stars created, can be constrained by a theoretical model of
 \citet{1995ApJS..101..181W}.  This model suggests that $f_{Z*} = 0.02$
is an appropriate number.  The minimum dynamical time is set to be $T_{dyn,min} = 10^7$
~years to agree with timescales seen in nearby OB associations.

The other parameters, such as the overdensity threshold $\eta$, minimum star
 mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are not well constrained 
either theoretically or observationally.  Indeed, $m_{*min}$ is a purely
numerical parameter designed to keep the code from producing too many star 
particles, and thus has no observational or theoretical counterpart.  These
parameters have to be set by performing parameter studies and comparing to 
observations.  Unfortunately, the range of parameter space is large, and
the results may be degenerate for some combinations of these parameters.

% Specific example: Kravtsov
\subsubsection{Kravtsov}
\label{sec:starform_kravtsov}

% GB: Is this going to be in the code????
% GB: I assume not, so I am commenting it out.  Text to be used in another paper?

%Kravtsov's method of star formation is designed to reproduce
%the global Schmidt law of star 
%formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
%This algorithm is deliberately minimal, and is explicitly geared
%towards modeling star formation in a phenomenological way on kiloparsec
%scales.  Stars are assume to form with a characteristic gas timescale 
%$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
%and $\rho_*$ are the baryon gas and stellar densities, respectively.  
%This ``constant efficiency'' model on the scale of star formation 
%regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
%Star formation is only allowed to take place in very dense regions
%with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
%proper density threshold above which star formation is allowed to
%occur.  No other criteria are imposed.  Kravtsov's typical choices
%for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
%1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
%is derived from the observationally-determined normalization of the
%Schmidt law, and the density threshold is determined by observations
%of star forming regions on $\sim 100$ pc scales.  Note that the density
%threshold is in proper, not comoving, units.

%Algorithmically, the star formation events in Kravtsov's code
%are assumed to occur once every global time step 
%$\Delta t_0 \leq 10^7$.  In cells where star formation is
%determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
%collisionless ``star particle'' is assumed to form, with a 
%mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
%$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
%volume of the mesh cell.  These star particles are
%formed instantaneously, with all $m_*$ of gas being
%taken out of the cell and immediately deposited into the
%star particle.  This particle is then given the velocity
%of the gas in the cell, and thereafter treated as a 
%collisionless particle.  The \enzo\ implementation
%of this algorithm is similar, except that instead of 
%forming stars only at the root grid time step, we allow 
%stars to form at the time step of the highest level of
%resolution at any particular point in space.  As can be seen
%from the equation for $m_*$ above, this can result in very 
%small stellar masses.  To avoid memory and processor time
%issues related to having very large numbers of star particles
%we impose a threshold mass $M_{*,min}$ such that a star
%particle only forms if $m_* \geq M_{*,min}$.  An
%appropriate choice of $M_{*,min}$ does not significantly 
%change the star overall star formation history of a 
%simulation, though it may delay the onset of star formation 
%in a given cell relative to a simulation without a particle mass
%threshold.

%Each ``star particle'' is assumed to represent an ensemble
%of stars and is treated as a single-age stellar population.
%Kravtsov assumes that the IMF is described by a
%Miller \& Scalo functional form with stellar masses between
%$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
%IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
%thermal energy and a mass $f_z M_*$ of metals into the
%cell in which they form without delay, with
%$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
%deposition of metals).  The definition of $f_z$ is a rough
%approximation of the results of 
%\citet{1995ApJS..101..181W}.

%Kravtsov reports that simulations with this algorithm
%reliably reproduce the star formation rate-gas surface
%density relation of the Schmidt law, and are not particularly
%sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
%He also notes that this is surprisingly insensitive to
%the presence or absence of feedback and details of the cooling
%and heating properties of the gas, which suggests that the
%global star formation rate is determined by gravitationally-
%driven supersonic turbulence (on large scales)
%rather than stellar feedback or thermal instabilities on small 
%scales.

% Specific example: H2-regulated
\subsubsection{H2-regulated}
\label{sec:starform_H2reg}

% Specific example: Pop III
\subsubsection{Population III}
\label{sec:starform_pop3}

%\subsubsection{The Kravtsov star formation and feedback algorithm}