### 2013-03-15 numerical-cooling.tex

\subsection{Radiative Cooling and Heating}
\label{sec.num.cooling}

\enzo has multiple methods for computing the energy change from
radiative cooling and heating.  All of them assume the gas is totally
optically thin.  Below we describe the methods for computing the
cooling rates from metal-free and metal-enriched gas.  Sample cooling
curves for each of \enzo's primary cooling methods are shown in Figure
\ref{fig.cooling_rate}.

\subsubsection{Primordial Cooling}

As discussed in \S\ref{sec.ov.chem}, the set of reactions that
characterize a metal-free gas is simple enough to be computed in
non-equilibrium during the simulation.  Similarly, the radiative
cooling of metal-free gas is solved by directly computing the cooling
and heating rates from the following individual processes for atomic
H and He: collisional excitation and ionization, recombination,
free-free emission, Compton scattering off of the cosmic microwave
background (CMB), and photo-heating from a UV metagalactic background.
If the H$_{2}$ chemistry network is enabled, the following H$_{2}$
cooling processes are also considered: ro-vibrational transitions 
\citep{1998A&A...335..403G}, heating and cooling from molecular
formation and destruction \citep{2009Sci...325..601T}, and 
collision-induced emission \citep{2004MNRAS.348.1019R}.  If Deuterium
chemistry is enabled, then rotational transitions of HD
\citep{1998A&A...335..403G} are treated as well.  The radiative
cooling calculation is coupled to the update of the chemistry network
such that they both occur within the same subcycling loop.  This is
necessary in regimes where rapid cooling or change in the ionization
state occur.  In
addition to the subcycle timestepping contraints mentioned in
\S\ref{sec.ov.chem}, the subcycle timestep is also not permitted to
exceed 10\% of the cooling time, $e/\dot{e}$.  A metagalactic
background affects the gas through both photo-heating and
photo-ionization.  These are treated by including redshift-dependent 
photo-ionization and photo-heating rate terms in the chemistry and
cooling equations for HI, HeI, and HeII.  The black curves in Figure
\ref{fig.cooling_rate} show cooling rates for a metal-free gas with
number density $n_{H} = 10^{-4}$ cm$^{-3}$ both with and without a
radiation background.  More detail on the specific UV backgrounds
present in \enzo is presented below in
\S\ref{sec.num.rad-homogeneous}.

\subsubsection{Metal Cooling}

A proper treatment of the cooling from metals is significantly more
challenging due to the large number of chemical reactions and energy
transitions that must be taken into account for each element.  Because
of this, most metal cooling methods employ significant assumptions in
order to seek out a balance between accuracy and speed.  There are two
primary metal cooling methods available in \enzo.  The simpler of the
two uses the analytic cooling function of \citet{SW87}, which assumes
a fully ionized gas with a constant metallicity of 0.5 Z$_{\odot}$.
The cooling rate produced by this cooling function is shown by the red 
curve in Figure \ref{fig.cooling_rate}.

A more sophisticated method makes use of multidimensional cooling and
heating rate tables computed with the photo-ionization code,
\texttt{Cloudy} \citep{1998PASP..110..761F}.  This method, detailed in
\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}, works by
using \texttt{Cloudy} to compute the cooling and heating rates from
the metal species only.  The primary assumption made is that of
ionization equilibrium.  The tables can have up to five dimensions,
where those dimensions are density, metallicity, electron fraction,
redshift (for an evolving metagalactic UV backgroun), and
temperature.  Tables can be created for any abundance pattern for
elements up to atomic number 30 (Zn) and for any incident radiation
field.  The cooling from H and He is calculated by the non-equilibrium
primordial chemistry and cooling solver.  The contribution from metals
to the cooling is computed within the same subcycling loop as the
coupled primordial chemistry and cooling solver.  The blue curves in
Figure \ref{fig.cooling_rate} show cooling rates calculated with this
method for a gas with number density $n_{H} = 10^{-4}$ cm$^{-3}$ and
metallicity of 0.5 Z$_{\odot}$.

\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{figures/cooling_rate.eps}
  \end{center}
  \caption{Radiative cooling rates from the various cooling methods
    available in \enzo.  The black curves show cooling rates from a gas
    with primordial composition using the non-equilibrium chemistry
    network.  The blue curves are for a gas with metallicity of 0.5
    $Z_{\odot}$ computed with the Cloudy cooling method.  The solid
    black and blue lines assume collisional processes only while the
    dashed lines include photo-ionization and photo-heating from a UV
    metagalactic background at $z = 0$ with a gas number density of
    10$^{-4}$ cm$^{-3}$.  The low temperature rates shown by the
    dashed lines indicate a net heating.  The red curve is the
    tabulated cooling function of \citet{SW87} which assumes a fully
    ionized gas with metallicity of 0.5 $Z_{\odot}$.}
  \label{fig.cooling_rate}
\end{figure}