### 2013-03-19 appendix.tex

\appendix
\section{Interpolation methods}
\label{app:interpolation}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ms"
%%% End: