### 2013-03-20 numerical-particle.tex

\subsection{N-body Dynamics}
\label{sec.ov.nbody}

% The code needs to be able to model collisionless matter.  One way
% would be to model the dark matter (or other collisionless
% particle-like objects, such as stars) as a second fluid in addition to
% the baryonic fluid and solve the collisionless Boltzmann equation,
% which follows the evolution of the fluid density in six-dimensional
% phase space.  However, this is computationally prohibitive owing to
% the large dimensionality of the problem, making this approach
% unappealing for the cosmological AMR code.
%GB: removed the above (good discussion but not really appropriate in this paper I think, or maybe at the beginning)

\enzo\ uses a particle-mesh N-body method to calculate the 
dynamics of collisionless systems \citep{Hockney88}.  This method 
follows trajectories of a representative sample of individual particles 
and is much more efficient than a direct solution of the Boltzmann 
equation in essentially all astrophysical situations. 
As described earlier, the gravitational potential is computed by solving the elliptic 
Poisson's equation (Eq.~\ref{eq:potential}), and differencing the potential
to find accelerations, which are then interpolated back to particles.  This
acceleration is time-centered (because the underlying gravitating mass
field is approximately time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$
for each particle.  These are used to update the particle positions and velocities
starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a standard drift-kick-drift technique:
\begin{eqnarray}
\label{eqn.driftkick}
\myvec{x}^{n+1/2} & = & \myvec{x}^n + \frac{\Delta t}{2 a^n} \myvec{v}^{n} \nonumber \\
\myvec{v}^{n+1} & = & \myvec{v}^n \left(1 - \frac{\dot{a}^{n+1/2}}{a^{n+1/2}}\right) + \frac{\Delta t}{ a^{n+1/2}} \myvec{a}^{n+1/2} \\
\myvec{x}^{n+1} & = & \myvec{x}^{n+1/2} + \frac{\Delta t}{2 a^{n+1}} \myvec{v}^{n+1} \nonumber
\end{eqnarray}

Particles are stored in the most highly refined grid patch at the point in space where they exist, and particles which move out of a subgrid patch are sent to the grid patch covering the adjacent volume with the finest spatial resolution, which may be of the same spatial resolution, coarser, or finer than the grid patch that the particles are moved from.  This takes place in a communication process at the end of each timestep on a level.