### 2013-03-25 conclusions.tex
finishing up conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

In this paper, we have presented the algorithms underlying \enzo, an
open-source adaptive mesh refinement code designed for
self-gravitating compressible fluid dynamics, including the effects of
magnetic fields, radiation transport, and a variety of microphysical
and subgrid processes.  In addition, we have described the \enzo\ code
development process, have shown the outputs of some representative
test problems, and have provided information about \enzo's performance
and parallel scaling on recent supercomputing platforms.  The \enzo\
code, its test suite, and all of the scripts used to generate plots
and figures for this paper are open source and are available at the
\enzo\ website, http://enzo-project.org.  Furthermore, the yt toolkit,
which is designed to analyze \enzo\ data (as well as data from a wide
variety of other simulation tools), can be found at its website,
http://yt-project.org.  Both of these codes have active user and
developer communities, extensive documentation and user suppport, and
strong mechanisms for users to contribute their changes and fixes to
the codebase.

The developers of the \enzo\ code are currently working on several
projects that will extend the functionality, scalability, or overall
performance of the code in the near future.  Projects that will appear
in forthcoming releases of the \enzo\ code include:

\begin{itemize}
\item The creation of a hybrid-parallel version of \enzo, combining
MPI for communication between nodes of a supercomputer and OpenMP for
thread-based parallelism within a node.  This will reduce on-node
memory usage and overall scaling behavior.
\item The restructuring of \enzo's treatment of particles to
accommodate a wider range of ``active'' particles that can easily
interact with each other and with multiple grids, and include sink,
source, and particle creating, splitting, and merging functionality.
\item A new HYPRE-based AMR gravity solver that is faster, more
accurate, and more scalable than the current multigrid solver.
\item New infrastructure for problem initialization, enabling
  users to more quickly and easily create new types of simulations.
\end{itemize}

In the longer term, it seems likely that trends in supercomputing
hardware will require substantial modifications to \enzo's core
infrastructure, and very possibly some of the core algorithsms.  More
specifically, these trends involve the usage of specialized
large-core-count, vectorized computing units such as graphics
processuing units (GPUs) or chips like the Intel Xeon Phi, as well as
precipitously decreasing amounts of RAM per computing core.  The
former trend means that the amount of processing power per compute
node will continue to trend upward, likely much faster than the
bandwidth between nodes, and will require tremendous reduction in
inter-node (and possibly inter-CPU) communication in order to maintain
code scalability.  Also, much of the current code will need to be
rewritten to take advantage of the vector nature of these CPUs, making
assumptions that are quite unlike those made in much of the current
codebase.  The latter trend means that duplication of data -- for
example, the grid hierarchy -- must be effectively eliminated to save
memory, and all inter-core and inter-node communication must be
carefully thought through to minimize the amount of data moved.  An
additional challenge as one goes to core counts in the tens to
hundreds of millions (or more) is that the reliability of individual
computing elements will become much more of an issue, requiring
robustness to hardware failure to be built into the code at some
level.  Furthermore, we are nearing the physical limits of transistor
speed and latency~\citep{feynman1999feynman}, meaning that simple
hardware improvements will not make these challenges disappear, and
careful thought (and the rewriting of a great deal of code) must take
place!  These challenges are not unique to the \enzo\ code, and in
fact are faced by effectively all applications that wish to take
advantage of new computational architectures.  Thus, we suspect that
\enzo\ (or a code that has the capabilities of \enzo, from a user's
point of view) will continue to be usable at the largest scales on
such machines.