### 2013-04-22 ms.tex
Revised abstract.

%\documentclass[10pt,preprint]{aastex}  % for e-submission to ApJ
%\documentclass[12pt,preprint2]{aastex}  % for e-submission to ApJ - two column
\documentclass[onecolumn]{emulateapj}  % this makes everything look like ApJ

\usepackage{graphicx, natbib, color}

%%%% PUT NEW COMMANDS AND DEFINITIONS HERE %%%%
\input{local-commands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Paper Outline  (last updated January 18, 2013)
%
% 1.  Introduction
% 2.  Physical equations and overview
%   2.1 Physical equations
%   2.2 Overview of numerical methods
% 3.  Numerical methods
%   3.1   AMR
%   3.2.  Hydro: Godunov (PPM)
%   3.3   Hydro: Godunov (MUSCL)
%   3.4   Hydro: Finite difference (ZEUS)
%   3.5   MHD: Dedner
%   3.6   MHD: CT
%   3.7   Gravity
%   3.8   Particle Integration
%   3.9.  Chemistry
%   3.10 Cooling
%   3.11 Radiation: Homogeneous
%   3.12 Radiation: ray-tracing
%   3.13 Radiation: FLD
%   3.14 Thermal conduction
%   3.15 Star formation & Feedback
%   3.16 Timestep constraints
%   3.17 Analysis
% 4. Code tests
%   4.1 Verifying and Validating the Enzo code
%   4.2 Representative test problems
% 5. Parallelism 
%   5.1 Parallel strategy 
%   5.2 Parallel performance 
% 6. Code development methodology
% 7. Conclusions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{aa}  % correct formatting for ApJ style files

\begin{document}

\title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
\author{Authors}

%\altaffiltext{1}{ Department of Astronomy, Columbia University,
%New York, NY, 10027, U.S.A.  Email: gbryan@astro.columbia.edu }
%\altaffiltext{2}{Center for Astrophysics and Space Sciences,
%University of California at San Diego.
%Email: mnorman,dcollins,akritsuk,jbordner@cosmos.ucsd.edu}
%\altaffiltext{3}{Theoretical Astrophysics (T-6), Los Alamos National
%Laboratories.  Email:  bwoshea@lanl.gov}
%\altaffiltext{4}{San Diego Supercomputing Center, University of California
%at San Diego.  Email: harkness@sdsc.edu}
%\altaffiltext{5}{Kavli Institute for Particle Astrophysics and Cosmology,
% Stanford University.  Email:  tabel@stanford.edu}

\begin{abstract}
This paper describes the open-source \enzo\ code, which uses a
block-structured adaptive mesh refinement to provide high spatial and
temporal resolution for modeling astrophysical fluid flows.  The code
is Cartesian and can be run in 1, 2, and 3 dimensions, and supports a
large variety of physics, including hydrodynamics, ideal and non-ideal
magnetohydrodynamics, N-body dynamics (and, more broadly, self-gravity
of fluids and particles), primordial gas chemistry, optically-thin
radiative cooling of primordial and metal-enriched plasmas, radiation
transport, and cosmological expansion.  In addition to explaining the
algorithms implemented, we present solutions for a wide range of test
problems, demonstrate the code's parallel scaling properties, and
discuss the \enzo\ consortium's code development methodology.
\end{abstract}

\keywords{methods: numerical --- hydrodynamics}

\maketitle

%% Section: introduction
\input{introduction}

%% Section:  Overview
\input{overview}

%% Section: numerical methods
\section{Numerical Methods}
\label{sec.methods}

We now describe each of the numerical methods in detail.  The structure of this section mirrors the previous one, in which each method was briefly described; here we present the details.

\input{numerical-amr}
\input{numerical-ppm}
\input{numerical-muscl} % new
%\input{numerical-mhd-dedner} % new (combined with muscl)
\input{numerical-mhd-ct}  % new
\input{numerical-zeus}
\input{numerical-gravity}
\input{numerical-particle}
\input{numerical-chemistry}
\input{numerical-cooling} % new
\input{numerical-radiation-homogeneous} % new
\input{numerical-radiation-raytracing} % new
\input{numerical-radiation-fld} % new
\input{numerical-conduction} % new
\input{numerical-starformation}
\input{numerical-timestep}
\input{numerical-analysis} % new

%% Section: code tests
%%  this should probably be done with one file per test
\input{code-tests}

%% Section: parallelism and performance
\input{parallel}

%% Section: code development methodology
\input{development} % new

%% Section: conclusions
\input{conclusions}  % new

\input{appendix}  % new

\bibliographystyle{apj}
\bibliography{apj-jour,ms}  % looks in ms.bib for bibliography info

\end{document}