### 2013-05-21 tests-pointsourcegravity.tex

\subsubsection{Point source gravity test}
\label{sec.test.gravitypointsource}

This test is a simple test of the ability of the Poisson solver to correctly reproduce the acceleration around a point source without any dynamics.  We place a single particle in the center of the unit domain with root grid dimensions $32^3$ and add a static AMR grid with refinement factor two that covers the central $1/8^3$ of the domain (i.e. so the refined grid is only $8^3$).  We then place 5000 particles throughout the domain distributed uniformly in angle and with the logarithm of the radius (from the central point).  The measured accelerations are shown in Figure~\ref{fig.gravitytest}.  As is typical for Particle-Mesh based methods, the errors decrease at large distances and peak on roughly the grid scale.  The accelerations are relatively smooth across the top grid/fine grid boundary with the most noticeable impact being a drop in the amplitude of tangential errors on the fine grid.  The force law rolls over at about 1.6 fine cell widths (equivalent to 0.8 root grid widths in this figure), consistent with the CIC-deposition and interpolation used.

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/GravityTest.eps}
\caption{The radial and tangential accelerations measured around a point source placed in a static AMR grid.  The solid line shows the analytic radial profile.}
\label{fig.gravitytest}
\end{center}
\end{figure}