### 2013-05-30 appendix.tex
Added appendix

\appendix
\section{Interpolation methods}
\label{app:interpolation}

In this appendix, we provide details for the various interpolation methods available in the code.  We assume throughout that we are dealing with Cartesian coordinates and equally spaced cells, although we allow for arbitrary refinement factor $r$.

\vspace{0.5cm}\noindent
{\bf SecondOrderA}

This interpolation algorithm is generally second-order but has monotonicity constraints as described below.

In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at $x_0$ and width $\Delta x$. We first linearly interpolate the parent values to the cell edges: $Q_{-1/2} = (Q_0 + Q_{-1})/2$, and similarly for $Q_{1/2} = (Q_0 + Q_1)/2$, and then compute a monotonic slope: $\Delta Q_0 = \minmod{Q_{1/2} - Q_0}{Q_0 - Q_{-1/2}}$ where
\begin{equation}
\minmod{a}{b} = \left\{ \begin{array}{ll}
0 & {\rm if} \quad ab < 0 \\
\min{ \left( | a |, | b | \right) } {\rm sign} (a) & {\rm otherwise}
\end{array}\right.
\end{equation}
This slope is then used to compute the interpolated subgrid values.  Defining $q_i$ to be the subgrid values at cell centers for the $r$ cells corresponding to parent cell $Q_0$ (for refinement factor $r$), we can write:
\begin{equation}
q_i = Q_0 + \frac{i+(1-r)/2}{r} f_x
\end{equation}
where $f_x = 2 \Delta Q_0$ (this notation is used for consistency with the 2 and 3 dimensional cases).

In two dimensions, the procedure is very similar in that we linearly interpolate parent values to the four cell corners: $Q_{-1/2, -1/2}$, $Q_{-1/2, 1/2}$, $Q_{1/2, -1/2}$, $Q_{1/2, 1/2}$ (by averaging parent cell-centered parent values).  We then compute monotonic slopes across the two diagonals, since for a linear function, extrema occur at corners:
\begin{eqnarray}
\Delta Q_0 & = & \minmod{Q_{1/2, 1/2} - Q_0}{Q_0 - Q_{-1/2, -1/2}} \\
\Delta Q_1 & = & \minmod{Q_{-1/2, 1/2} - Q_0}{Q_0 - Q_{1/2, -1/2}}
\end{eqnarray}
We then translate these slopes into the grid axes with $f_x = \Delta Q_0 -  \Delta Q_1$ and $f_y = \Delta Q_0 + \Delta Q_1$ so that the interpolation itself can be written simply as
\begin{equation}
q_{i,j} = Q_0 + \frac{i+(1-r)/2}{r} f_x + \frac{j+(1-r)/2}{r} f_y
\end{equation}

Finally, we write down the three-dimensional version -- unfortunately, here the number of monotonicity constraints is four (the 4 diagonals across the 8 opposing corners of the cube), while the number of slopes is three, so the problem is over constrained.  Somewhat arbitrarily, we adopt the following procedure.  As before, we compute $\Delta Q_0$, $\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$ with the minmod limiter across the 4 diagonals.  We then define
\begin{equation}
s =\frac{\Delta Q_1 + \Delta Q_2 + \Delta Q_3}{\Delta Q_0}
\end{equation}
which is the value of $\Delta Q_0$ that the other slopes ($\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$) imply, normalized by the desired value of $\Delta Q_0$ itself.  If $0 < s < 1$, then no adjustment needs to be made, as the monotonicity constraints are met.  If these equalities are not met, then, if $s<0$ we define $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 < 0$, and 0 otherwise (if $s > 1$, this is reversed, so $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 > 0$, and 0 otherwise).  These weights are used to determine which slopes to modify to match the $\Delta Q_0$ constraint.  We then compute the amount of adjustment required:
\begin{equation}
f = - \frac{\Delta Q_0 s^\prime + \sum_n (1-\chi_n) \Delta Q_n}{\sum \chi_n \Delta Q_n + \epsilon}
\end{equation}
where $s^\prime = \min(\max(s,0), 1)$ and $\epsilon$ is a small number to prevent numerical errors. We then use this adjustment fraction to compute the new adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match the $\Delta Q_0$ constraint as closely as possible with
\begin{equation}
\Delta Q_n = f \chi_n \Delta Q_n + (1-\chi_n) \Delta Q_n
\end{equation}
Finally, these are converted to the grid axes with $f_x = \Delta Q_2 + \Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 + \Delta Q_2$ and then we do the interpolation itself with,
\begin{equation}
q_{i,j,k} = Q_0 + \frac{i+(1-r)/2}{r} f_x + \frac{j+(1-r)/2}{r} f_y + \frac{k+(1-r)/2}{r} f_z
\end{equation}

% --------------

\vspace{0.3cm}\noindent
{\bf SecondOrderB}

We also provide a variant on the above procedure, with two changes.  The first is that the slopes across the diagonals ($\Delta Q_0$, etc.) are computed directly rather than with the minmod limiter (e.g. $\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).  To ensure positivity in the resulting interpolation when applied to positive conserved quantities, the slopes are limited so that the smallest corner value is 0.2 the cell center value.  The same procedure as described above is then applied to turn these four slopes into a linear interpolation.

% --------------

\vspace{0.3cm}\noindent
{\bf SecondOrderC}

This is a completely different second-order interpolation scheme, which is based on Cloud-In-Cell (CIC) interpolation \citep{Hockney88}.  In one dimension, we define the parent values as $Q_0$, and $Q_{+1}$, where the left parent cell has a cell center at $x_0$ and width $\Delta x$.  Then, the interpolated value for a subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is given by:
\begin{equation}
q_i =  \frac{2r - 1 - 2i}{2r} Q_0 + \frac{1+2i}{2r} Q_{+1}
\end{equation}
The extension to two and three dimensions is straightforward, with weights in the other dimensions computed in the same way and then multiplied to get a total of 4 and 8 weights for the 2 and 3 dimensional cases, respectively.  This scheme is monotonicity preserving, but is not conservative.

% --------------

\vspace{0.3cm}\noindent
{\bf ThirdOrderA}

This interpolation method provides a third-order accuracy based on the Triangular Shaped Cloud (TSC) methodology \citep{Hockney88}.  As usual, in one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at $x_0$ and width $\Delta x$.  Then, the interpolated value for a subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is given by:
\begin{equation}
q_i = a_i  Q_{-1} + b_i Q_0 + c_i Q_{+1}
\end{equation}
and the weights are given by:
\begin{equation}
a(i) =  \frac{(r-i)^3 - (r-i-1)^3}{6r^3}; \qquad c(i) = \frac{3i^2 + 3i + 1}{6r^3}
\end{equation}
and $b(i) = 1/r - a(i) - c(i)$.  The extension to two and three dimensions is straightforward with the weights computed in the same way as for the one-dimensional case but then multiplied to determine the 9 and 27 factors necessary for the 2 and 3 dimensional cases, respectively.

One problem with this interpolation technique is that it is not conservative.  In particular, the sum of the interpolated subgrid values:
\begin{equation}
\tilde{Q_0} = \sum_{i=0}^{r-1} q_i
\end{equation}
is not, in general, equal to $Q_0$.  We can retain conservation by adding the factor $(Q_0 - \tilde{Q_0})/r$ to the interpolated values (or by multiplying the interpolated values by the ratio $Q_0/\tilde{Q_0}$).  Unfortunately, the result of this procedure does not preserve monotonicity -- it can introduce local minima and maxima at at parent cell boundaries.  We can then attempt to correct that by taking a weighted average between the interpolated values on the subgrid and the parent value with weights computed such that the interpolated values at the edge of the parent cell are not local maxima compared to the interpolated values in the neighboring parent cell.

% --------------

\vspace{0.3cm}\noindent
{\bf FirstOrderA}

Finally, for completeness, we include a first-order accurate piecewise constant interpolator, for which, using the same definitions in the previous case, we take $q_i = Q_0$, for $i = 0$ to $r-1$.

\vspace{1cm}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ms"
%%% End: