### 2013-06-07 development.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code development methodology}
\label{sec.development}

Over time, \enzo's development has followed a trajectory toward
openness.  Started as the graduate research project of Greg Bryan at
the University of Illinois, it was subsequently stewarded by the
Laboratory for Computational Astrophysics (LCA) at the University of
California at San Diego, and has transitioned to a distributed,
completely open, and community-driven project.  Initially, \enzo\ was
versioned using a series of ``snapshots'' of the code base, usually
hand-created by the individuals doing development work.  These were
distributed to collaborators and colleagues, but the central ``trunk''
of development was updated primarily by a single person: while patches
and technology were accepted from external developers, the relatively
small number of individuals using the code resulted in a strong
centralization of development.

As the stewardship of the code passed to the LCA, the code was
released first to ``friendly users'' and then as an open source
release.  However, while the code was made available with
documentation, technology developed in the broader community of users
was typically not re-integrated.  This led to a wide dispersal of
development, largely independent, by individuals who downloaded and
used the version of the code developed by the LCA.

Following the first public, open source release of \enzo, the code was
migrated to the Subversion version control system.  This is a
centralized version control system, and the ``stable'' \enzo\ source
code was made globally readable following the \enzo\ 1.5 release.
Access to the primary development tree required a password and login
for each user, and providing upstream changes either required this
password and the granting of write access or a sequence of patches and
manually-created diffs (much like the original development system).
The technical friction of manually contributing patches and
modifications, combined with Subversion's difficulty with tracking
merges, resulted in further fragmentation of the code base.

A version developed at Penn State and Stanford forked from a version
prior to the LCA version and was the one in which MHD with Dedener
divergence cleaning, the MUSCL hydro solvers, the ray tracing
radiative transfer module, relativistic fluid dynamics, the shearing
box boundaries and updates to the multi-species chemistry were
included. This version was the one that was eventually merged using
the distributed version control system (DVCS) Mercurial
(\url{http://mercurial.selenic.com/}) into a branch of the code known
as \texttt{week-of-code}, so named after the in-person development
sprint at KIPAC in June 2009 at which it was created.  The
fundamental, and transformative, distinction between the previous
\textit{centralized} version control system and mercurial's
\textit{distributed} version control system is the elimination of
gatekeepers.  While there still exists a canonical, central location
where stable and development versions of \enzo\ can be obtained,
changesets and versions can be exchanged between peers without the
intervention of designated gatekeepers.  This has the direct effect of
enabling local development to be versioned and its provenance ensured,
while still retaining the ability to benefit from ``upstream''
development.  An important, even crucial, side effect is that the
technology used for local versioning provides mechanisms for easily
submitting locally-developed modifications to the community source
location.  Mercurial internally represents all changes as nodes in a
directed, acyclic graph (DAG), which results in the natural ability to
more consistently and easily manage merging development streams.

Currently, \enzo\ is developed using the hosted source control
platform Bitbucket (\url{http://bitbucket.org/}) at
\url{http://bitbucket.org/enzo/}.  There are two mailing lists, one
for usage-focused questions and discussion, and another for
development discussion.  Both of these lists are open and publicly
archived.  Bitbucket provides mechanisms for inspecting source code,
hosting branches and forks of the primary source, and for code review.
All proposed source code changes for \enzo\ are subjected to a peer
review process, where experienced developers read, inspect, test, and
provide feedback on source code changes.  All developers, including
long-time \enzo\ contributors and developers, are subject to this
process before their code is included in the primary \enzo\
repository.  By using a remote, hosted system, \enzo\ is now
\textit{completely} open to contributions from the community.
Individuals, who may or may not consider themselves \enzo\ developers,
are free to ``fork'' the \enzo\ code base, develop changes (signed
with their own name), and submit them for review and inclusion.  In
contrast to the centralized, gatekeeper-focused technology used
previously, this enables anyone to contribute changes to be evaluated
for inclusion in the \enzo\ codebase.

While peer review is able to catch many bugs and problems with source
code changes, \enzo\ is also subject to ``answer'' testing (described
in detail in Section~\ref{sec.tests.suite}).  We have created a set of
parameter files and problem types that exercise the underlying
machinery of \enzo.  These ``test problems'' have affiliated
``tests,'' which consist of scripts that use \texttt{yt}
\citep{2011ApJS..192....9T} to produce results such as mass
distribution, projections, profiles and so on.  The testing
infrastructure then evaluates whether the variation in the new results
compared to a ``gold standard'' of results has exceeded an acceptable
threshold, typically set to roughly the level of roundoff error in
single-precision floating point arithmetic.  Optionally, for those
test problems that are deemed unsafe to change to any precision, the
tests also produce hashes of the outputs; these hashes will only
remain unchanged in the event of bitwise identicality between results.
The results of the gold standard are versioned and stored in Amazon
S3, enabling remote testing to proceed.  While the testing process --
building, running, analyzing and comparing -- is not yet automated
against incoming pull requests, we hope to deploy that functionality
in the future.  The primary challenge is that of compute time; the
tests are organized into multiple categories, including by the
expected run time, but the full suite of tests can take several days
to run.