### 2013-06-08 ms.tex
Added author list

%\documentclass[10pt,preprint]{aastex}  % for e-submission to ApJ
%\documentclass[12pt,preprint2]{aastex}  % for e-submission to ApJ - two column
\documentclass[onecolumn,iop]{emulateapj}  % this makes everything look like ApJ

\usepackage{graphicx, natbib, color, bm}

%%%% PUT NEW COMMANDS AND DEFINITIONS HERE %%%%
\input{local-commands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Paper Outline  (last updated May 17, 2013)
%
% 1.  Introduction
% 2.  Physical equations and overview of numerical methodology
%   2.1 Physical equations
%   2.2 Overview of numerical methods
% 3.  The structured AMR method
% 4.  Fluid methods
%   4.1 PPM
%   4.2 MUSCL
%   4.3 MHD-CT
%   4.4 ZEUS 
% 5.  Gravity and N-body
%   5.1 Gravity
%   5.2 N-body dynamics
% 6. Microphysics
%   6.1  Chemistry
%   6.2  Radiative Cooling and heating
% 7. Radiation
%   7.1 Radiation: Homogeneous
%   7.2 Radiation: ray-tracing
%   7.3 Radiation: FLD
% 8. Other physics
%   8.1 Thermal conduction
%   8.2 Star formation & Feedback
% 9. Timestepping
% 10.  Analysis
% 11. Code tests
%   11.1 Verifying and Validating the Enzo code
%   11.2 Representative test problems
% 12. Parallelism 
%   12.1 Parallel strategy 
%   12.2 Parallel performance 
% 13. Code development methodology
% 14. Conclusions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{aa}  % correct formatting for ApJ style files

\begin{document}

\title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
\author{The Enzo Collaboration:}
\author{Greg L. Bryan\altaffilmark{1}}
\author{Michael L. Norman\altaffilmark{2,3}}
\author{Brian W. O'Shea\altaffilmark{4,5}}
\author{Tom Abel\altaffilmark{6, 19}}
\author{John H. Wise\altaffilmark{7}}
\author{Matthew J. Turk\altaffilmark{1}}
\author{Daniel R. Reynolds\altaffilmark{8}}
\author{David C. Collins\altaffilmark{9}}
\author{Peng Wang\altaffilmark{6}}
\author{Samuel W. Skillman\altaffilmark{10,11}}
\author{Britton Smith\altaffilmark{4}}
\author{Robert P. Harkness\altaffilmark{12}}
\author{James Bordner\altaffilmark{2}}
\author{Ji-hoon Kim\altaffilmark{13}}
\author{Michael Kuhlen\altaffilmark{14,15}}
\author{Hao Xu\altaffilmark{2}}
\author{Nathan Goldbaum\altaffilmark{15}}
\author{Cameron Hummels\altaffilmark{16}}
\author{Alexei Kritsuk\altaffilmark{2}}
\author{Elizabeth Tasker\altaffilmark{17}}
\author{Stephen Skory\altaffilmark{2}}
\author{Christine Simpson\altaffilmark{1}}
\author{Oliver Hahn\altaffilmark{18}}
\author{Jeffrey S. Oishi\altaffilmark{6}}
\author{Geoffrey C So\altaffilmark{2}}
\author{Fen Zhao\altaffilmark{19}}
\author{Renyue Cen\altaffilmark{20}}
\author{Yuan Li\altaffilmark{1}}

\altaffiltext{1}{Columbia University, Department of Astronomy, New York, NY, 10025, USA}
\altaffiltext{2}{CASS, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0424, USA}
\altaffiltext{3}{SDSC, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0505, USA}
\altaffiltext{4}{Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}
\altaffiltext{5}{Lyman Briggs College and Institute for Cyber-Enabled Research, Michigan State University, East Lansing, MI 48824, USA}
\altaffiltext{6}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, Menlo Park, CA 94025, USA}
\altaffiltext{7}{Center for Relativistic Astrophysics, School of Physics, Georgia Institute of Technology, 837 State Street, Atlanta, GA, USA}
\altaffiltext{8}{Southern Methodist University, 6425 Boaz Ln, Dallas, TX 75205, USA}
\altaffiltext{9}{Theoretical Division, Los Alamos National Laboratory, Los Alamos, NM 87544}
\altaffiltext{10}{Center for Astrophysics and Space Astronomy, Department of Astrophysical and Planetary Science, University of Colorado, Boulder, CO 80309, USA}
\altaffiltext{11}{DOE Computational Science Graduate Fellow}
\altaffiltext{12}{NICS, Oak Ridge National Laboratory, 1 Bethel Valley Rd, Oak Ridge, TN 37831, USA}
\altaffiltext{13}{Department of Astronomy and Astrophysics, University of California, Santa Cruz, CA 95064, USA}
\altaffiltext{14}{Theoretical Astrophysics Center, University of California Berkeley, Hearst Field Annex, Berkeley, CA 94720, USA}
\altaffiltext{15}{Department of Astronomy and Astrophysics, University of California, 1156 High Street, Santa Cruz, CA 95064, USA}
\altaffiltext{16}{Department of Astronomy/Steward Observatory, University of Arizona, 933 N. Cherry Ave., Tucson, AZ, 85721, USA}
\altaffiltext{17}{Physics Department, Faculty of Science, Hokkaido University, Kita-10 Nishi 8, Kita-ku, Sapporo 060-0810, Japan}
\altaffiltext{18}{ETH Zurich Institute for Astronomy. CH-8093 Zurich. Switzerland}
\altaffiltext{19}{Department of Physics, Stanford University, Stanford, CA 94305, USA}
\altaffiltext{20}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}

\begin{abstract}
This paper describes the open-source code \enzo, which uses a
block-structured adaptive mesh refinement to provide high spatial and
temporal resolution for modeling astrophysical fluid flows.  The code
is Cartesian and can be run in 1, 2, and 3 dimensions, and supports a
large variety of physics, including hydrodynamics, ideal and non-ideal
magnetohydrodynamics, N-body dynamics (and, more broadly, self-gravity
of fluids and particles), primordial gas chemistry, optically-thin
radiative cooling of primordial and metal-enriched plasmas, radiation
transport, and cosmological expansion.  In addition to explaining the
algorithms implemented, we present solutions for a wide range of test
problems, demonstrate the code's parallel scaling properties, and
discuss the \enzo\ collaboration's code development methodology.
\end{abstract}

\keywords{methods: numerical --- hydrodynamics}

\maketitle

%% Section: introduction
\input{introduction}

%% Section:  Overview
\input{overview}

%% Section: numerical methods
% \section{Numerical Methods}
% \label{sec.methods}

% We now describe each of the numerical methods in detail.  
\input{numerical-amr}

\section{Fluid methods}
\label{sec.fluids}

\input{numerical-ppm}
\input{numerical-mhd-dedner}  % includes MUSCL and Dedner
\input{numerical-mhd-ct}  
\input{numerical-zeus}

\section{Gravity and N-body}
\label{sec.allgrav}
\input{numerical-gravity}
\input{numerical-particle}

\section{Microphysics}
\label{sec.microphysics}
\input{numerical-chemistry}
\input{numerical-cooling}

\section{Radiation}
\label{sec.allrad}
\input{numerical-radiation-homogeneous} 
\input{numerical-radiation-raytracing} 
\input{numerical-radiation-fld}

\section{Other physics}
\label{sec.otherphys}
\input{numerical-conduction} 
\input{numerical-starformation}

% this is its own section
\input{numerical-timestep}

% this is its own section
\input{numerical-analysis}

%% Section: code tests
%%  this should probably be done with one file per test
\input{code-tests}

%% Section: parallelism and performance
\input{parallel}

%% Section: code development methodology
\input{development}

%% Section: conclusions
\input{conclusions}

\input{acknowledgments}

\input{appendix}

\bibliographystyle{apj}
\bibliography{apj-jour,ms}  % looks in ms.bib for bibliography info

\end{document}