### 2013-06-12 numerical-radiation-raytracing.tex

\subsection{Radiation transport: ray tracing}
\label{sec.num.raytracing}

Stars and black holes strongly affect their surroundings through radiation.
Radiation transport is a well-studied problem; however, its treatment in
multidimensional calculations is difficult because of the dependence on seven
variables -- three spatial, two angular, frequency, and time.  The non-local
nature of the thermal and hydrodynamical response to radiation sources further
adds to the difficulty.  Here we briefly describe \enzo's ray tracing
implementation \moray, which is presented in full detail in
\citet{Wise11_Moray} with seven code tests and six applications.

%The radiative transfer equation in comoving coordinates
%\citep{Gnedin97} reads
%%
%\begin{equation}
%  \label{eq:rteqn}
%  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
%  \frac{\hat{n} \cdot \grad I_\nu}{\bar{a}} -
%  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
%  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
%\end{equation}
%%
%Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
%specific intensity in units of energy per time $t$ per solid angle per
%unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
%constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
%ratio of scale factors at the current time and time of emission.  The
%second term represents the propagation of radiation, where the factor
%$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
%both the cosmological redshift and dilution of radiation.  On the
%right hand side, the first term considers the absorption coefficient
%$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
%$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
%that includes any point sources of radiation or diffuse radiation.
%GB: this was moved to the equations section.

We solve the radiative transfer equation in comoving coordinates (given by Equation~\ref{eq:rteqn}).
We can make some appropriate approximations to reduce the complexity
of this equation in order to include radiation transport
in numerical calculations.  Typically timesteps in dynamic
calculations are small enough so that $\Delta a/a \ll 1$, therefore
$a_{em}/a \approx 1$ in any given timestep, reducing the second term to
$\hat{n} \partial I_\nu/\partial \mathbf{x}$.  To determine the
importance of the third term, we evaluate the ratio of the third term
to the second term.  This is $HL/c$, where $L$ is the simulation box
length.  If this ratio is $\ll 1$, we can ignore the third term.  For
example at $z=5$, this ratio is 0.1 when $L = c/H(z=5)$ = 53 proper
Mpc.  In large boxes where the light crossing time is comparable to
the Hubble time, then it becomes important to consider cosmological
redshifting and dilution of the radiation.  Thus equation
(\ref{eq:rteqn}) reduces to the non-cosmological form in this local
approximation,
%
\begin{equation}
  \frac{1}{c} \frac{\partial I_\nu}{\partial t} + 
  \hat{n} \frac{\partial I_\nu}{\partial \mathbf{x}} =
  -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
We choose to represent the source term $j_\nu$ as point sources of
radiation (e.g. stars, quasars) that emit radial rays that are
propagated along the direction $\hat{n}$.

Ray tracing is an accurate method to propagate radiation from point
sources on a computational grid as long as there are a sufficient
number of rays passing through each cell.  Along a ray, the radiation
transfer equation reduces to
%
\begin{equation}
\label{eqn:rtray}
\frac{1}{c} \frac{\partial P}{\partial t} + \frac{\partial P}{\partial
  r} = -\kappa P,
\end{equation}
where $P$ is the photon number flux along the ray.  To sample the
radiation field at large radii, ray tracing requires at least $N_{ray}
= 4\pi R^2 / (\Delta x)^2$ rays to sample each cell with one ray,
where $R$ is the radius from the source to the cell and $\Delta x$ is
the cell width.  If one were to trace $N_{ray}$ rays out to $R$, the
cells at a smaller radius $r$ would be sampled by, on average,
$(r/R)^2$ rays, which is computationally wasteful because only a few
rays per cell  are required to provide an accurate calculation
of the radiation field (as we will show later).

We avoid this inefficiency by utilizing adaptive ray tracing
\citep{Abel02_RT}, which is based on Hierarchical Equal Area isoLatitude
Pixelation \citep[HEALPix;][]{HEALPix} and progressively splits rays when the sampling
becomes too coarse.  In this approach, the rays are
traced along normal directions of the centers of the HEALPix pixels that 
evenly divide a sphere into equal areas.  The rays are initialized at
each point source with the photon luminosity (photon s$^{-1}$) equally
spread across $N_{\rm pix} = 12 \times 4^l$ rays, where $l$ is the
initial HEALPix level.  We usually find $l$ = 0 or 1 is sufficient
because these coarse rays will usually be split before traversing
the first cell.

The rays are traced through the grid in a typical fashion
\citep[e.g.][]{Abel99_RT}, in which we calculate the next cell
boundary crossing.  The ray segment length crossing the cell is
%
\begin{equation}
  \label{eqn:trace}
  dr = R_0 - \min_{i=1 \rightarrow 3} \left[(x_{\rm cell,i} - x_{\rm src,i}) /
    \hat{n}_{\rm ray,i} \right],
\end{equation}
%
where $R_0$, $\hat{n}_{\rm ray}$, $x_{\rm cell,i}$, and $x_{\rm
  src,i}$ are the initial distance traveled by the ray, normal
direction of the ray, the next cell boundary crossing in the $i$-th
dimension, and the position of the point source that emitted the ray,
respectively.  However before the ray travels across the cell, we
evaluate the ratio of the face area $A_{\rm cell}$ of the current cell
and the solid angle $\Omega_{\rm ray}$ of the ray,
%
\begin{equation}
  \label{eqn:split}
  \Phi_c = \frac{A_{\rm cell}} {\Omega_{\rm ray}} = 
  \frac{N_{\rm pix} (\Delta x)^2} {4\pi R_0^2}.
\end{equation}
%
If $\Phi_c$ is less than a pre-determined value (usually $>3$), the
ray is split into 4 child rays.  The pixel numbers of the child rays
$p^\prime$ are given by the ``nested'' scheme of HEALPix at the next
level, i.e. $p^\prime = 4 \times p + [0,1,2,3]$, where $p$ is the
original pixel number.  The child rays (1) acquire the new normal
vectors of the pixels, (2) retain the same radius of the parent ray,
and (3) gets a quarter of the photon flux of the parent ray.
Afterward, the parent ray is discontinued.

A ray propagates and splits until at least one of the following conditions is met: (1) the photon has traveled $c
\times dt_P$, where $dt_P$ is the radiative transfer timestep; (2) its
photon flux is almost fully absorbed ($>99.9\%$) in a single cell,
which significantly reduces the computational time if the radiation
volume filling fraction is small; (3) the photon leaves the
computational domain with isolated boundary conditions; or (4) the
photon travels $\sqrt{3}$ of the simulation box length with periodic
boundary conditions.
In the first case, the photon is halted at that position and saved,
where it will be considered in the solution of $I_\nu$ at the next
timestep.  In the next timestep, the photon will encounter a different
hydrodynamical and ionization state, hence $\kappa$, in its path.
Furthermore any time variations of the luminosities will be retained
in the radiation field.  This is how this method retains the time
derivative of the radiative transfer equation.  The last restriction
prevents our method from considering sources external to the
computational domain. However, a uniform radiation background can be used
in conjunction with ray tracing that adds the background intensity to the
local radiation field.

The radiation field is calculated by integrating Equation
(\ref{eqn:rtray}) along each ray, which is done by considering the
discretization of the ray into segments.  In the following
description, we assume the rays are monochromatic for simplicity.  For
convenience, we express the integration in terms of optical depth
$\tau = \int \kappa(r,t) \; dr$, and for a ray segment
%
\begin{equation}
  \label{eqn:dtau}
  d\tau = \sigma_{\rm abs}(\nu) n_{\rm abs} dr.
\end{equation}
Here $\sigma_{\rm abs}$ and $n_{\rm abs}$ are the cross section and
number density of the absorbing medium, respectively.  In the static
case, equation (\ref{eqn:rtray}) has a simple exponential analytic
solution, and the photon flux of a ray is reduced by
%
\begin{equation}
  \label{eqn:flux}
  dP = P \times (1 - e^{-\tau})
\end{equation}
as it crosses a cell.  We equate the photo-ionization rate to the
absorption rate, resulting in photon conservation \citep{Abel99_RT,
  Mellema06}.  Thus the photo-ionization and
photo-heating rates associated with a single ray ($k_{\rm ph}$ and
$\Gamma_{\rm ph}$, respectively)  are
%
\begin{equation}
  \label{eqn:kph}
  k_{\rm ph} = \frac{P (1 - e^{-\tau})}{n_{\rm abs} \; V_{\rm cell} \; dt_P},
\end{equation}
\begin{equation}
  \label{eqn:gamma}
  \Gamma_{\rm ph} = k_{\rm ph} \; (E_{\rm ph} - E_i),
\end{equation}
where $V_{\rm cell}$ is the cell volume, $E_{\rm ph}$ is the photon
energy, and $E_i$ is the ionization energy of the absorbing material.
In each cell, the photo-ionization and photo-heating rates from each
ray in the calculation are summed. After the ray tracing is
complete, these rates are used as inputs to the solver described in
Section~\ref{sec.ov.chem} to update the ionization, chemical, and energy states
of the gas in each cell.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ms"
%%% End: