### 2013-06-24 table_of_variables_and_constants.tex

\begin{center}

%%% Use this to change the fontsize of the table
%\tabletypesize{\scriptsize}

%%% \LongTables command may not be necessary (allowed?) outside of
%%% emulateapj.cls...
\LongTables

\begin{deluxetable}{cl}
  \tablecaption{Summary of symbols used in this manuscript\label{tab:variables}}
  \tablehead{\colhead{Symbol} & \colhead{Description}}

  \startdata
  \multicolumn{2}{l}{\underline{Fluid Quantities}} \\[5pt]
  $E$ & comoving total fluid energy density \\
  $e$ & comoving thermal energy density \\
  $T$ & baryon temperature \\
  $S$ & fluid entropy \\
  $\rho$ & comoving baryonic density \\
  $\rho_{\!X}$ & comoving mass density of component $X$ (gas, dark matter, stars, metals, total, etc.) \\
  $t$ & time \\
  $\vecv$ & peculiar velocity \\
  $\vecB$ & comoving magnetic field strength \\
  $p$ & thermal pressure \\
  $p^*$ & total (thermal + magnetic) isotropic pressure \\
  $\gamma$ & ideal gas ratio of specific heats \\
  $\phi$ & gravitational potential \\
  $\Lambda$ & radiative/chemical cooling rate \\
  $\Gamma$ & radiative heating rate \\
  $Q$ & artificial viscosity stress tensor (only for ZEUS hydro) \\[3pt]

  \multicolumn{2}{l}{\underline{Code Quantities}} \\[5pt]
  $\Delta x_l$ & cell size on level $l$ \\
  $\Delta t_l$ & timestep on level $l$ \\
  $d$ & dimensionality (rank) of the simulation \\
  $q$ & conserved quantity \\
  $F$ & flux of quantity $q$ \\
  $r$ & refinement factor \\
  $\epsilon_l$ & refinement criterion parameter \\
  $J$ & number of cells per Jeans length (for Jeans length based refinement) \\[3pt]

  \multicolumn{2}{l}{\underline{Cosmology}} \\[5pt]
  $a$ & cosmological expansion factor \\
  $\Lambda_c$ & cosmological constant \\
  $\rho_0$ & mean comoving mass density (including both baryonic and dark matter) of the universe \\
  $p_0$ & comoving background pressure \\[3pt]

  \multicolumn{2}{l}{\underline{Chemistry}} \\[5pt]
  $n_i$ & comoving number density of species $i$ \\
  $k_{ij}$ & two-body reaction rate coefficient \\[3pt]

  \multicolumn{2}{l}{\underline{Ray-tracing Radiation Transfer}} \\[5pt]
  $\Gamma_j^{ph}$ & photo-ionization/-dissociation destruction/creation rates \\
  $I_\nu$ & specific radiation intensity (energy per time per solid angle per unit area per frequency) \\
  $\kappa_\nu$ & radiation absorption coefficient \\
  $j_\nu$ & radiation emission coefficient \\
  $P$ & photon number flux along ray \\
  $\sigma_{\rm abs}$ & absorption cross section \\
  $n_{\rm abs}$ & number density of absorbing medium \\
  $dt_P$ & radiative transfer timestep \\
  $k_{ph}$ & photo-ionization rate of a single ray \\
  $\Gamma_{ph}$ & photo-heating rate of a single ray \\[3pt]

  \multicolumn{2}{l}{\underline{Flux Limited Diffusion Radiation Transfer}} \\[5pt]
  $E_r$ & grey radiation energy density \\
  $e_c$ & internal energy correction due to photo-heating and chemical cooling \\
  $D$ & Larsen square-root flux-limiter \\
  $\kappa$ & total opacity \\
  $\eta$ & field of radiation sources \\[3pt]

  \multicolumn{2}{l}{\underline{Heat Conduction}} \\[5pt]
  $\fcond$ & thermal conduction heat flux \\
  $\kappa_{\rm cond}$ & thermal conduction coefficient \\
  $\kappa_{\rm sp}$ & Spitzer conductivity \\
  $f_{\rm sp}$ & fraction of Spitzer conductivity \\[3pt]

  \multicolumn{2}{l}{\underline{Star Formation And Feedback}} \\[5pt]
  $t_{\rm cool}$ & cooling time \\
  $t_{\rm dyn}$ & local dynamical time \\
  $t_{\rm form}$ & star particle formation time \\
  $m_{\rm b}$ & baryonic mass of cell \\
  $m_{\rm J}$ & Jeans mass of cell \\
  $m_*$ & star particle mass \\
  $m_{\rm *min}$ & minimum star particle mass \\
  $f_{\rm *eff}$ & star formation efficiency parameter \\
  $f_{\rm Zb}$ & metallicity fraction of baryon gas in cell \\
  $f_{\rm m*}$ & fraction of star particle's mass returned to cell \\
  $f_{\rm Z*}$ & fraction of star particle's metal mass returned to cell \\
  $f_{\rm SN}$ & fraction of stellar rest mass energy deposited as feedback \\
  $\Delta m_{\rm sf}$ & mass of stars formed in current timestep \\
  $E_{\rm add}$ & feedback energy added to cell \\
  $T_{\rm dyn,min}$ & user-defined minimum dynamical time \\
  $\eta$ & overdensity threshold for star formation \\
  $\rho_{\rm SF}$ & constant proper density threshold for star formation \\
  $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\[3pt]

  \multicolumn{2}{l}{\underline{Natural Constants and Values}} \\[5pt]
  $G$ & gravitational constant \\
  $H$ & Hubble constant \\
  $k_B$ & Boltzmann constant \\
  $M_\odot$ & mass of Sun \\
  $m_H$ & mass of hydrogen atom \\
  $m_e$ & mass of electron \\
  $c$ & speed of light \\
  $c_{\rm s}$ & sound speed \\
  $z$ & redshift \\
  \enddata

  %% \tablecomments{Optional comments}
\end{deluxetable}

\end{center}