### 2013-10-22 numerical-zeus.tex

\subsection{Hydrodynamics: The \zeus\ method}
\label{sec.hydro.zeus}

As an alternative to the previous Godunov methods, \enzo\ also
includes an implementation of the finite difference hydrodynamic
algorithm employed in the compressible magnetohydrodynamics code
\zeus\ \citep{Stone92a, Stone92b}.  Fluid transport is solved on a
Cartesian grid using the upwind, monotonic advection scheme of
\citet{1977JCoPh..23..276V} within a multistep (operator-split)
solution procedure that is fully explicit in time.  This method is
formally second-order accurate in space but first-order accurate in
time.

As discussed in the section describing the Piecewise Parabolic Method
(Section~\ref{sec.hydro.ppm}), operator-split methods break the
solution of the hydrodynamic equations into parts, with each part
representing a single term in the equations.  Each part is evaluated
successively using the results preceding it.  In this method, in
addition to operator-splitting the expansion terms (i.e., those terms
in Eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on
$\dot{a}$), we divide the remaining terms into \emph{source} and
\emph{transport} steps.  The terms to be solved in the source step are
those on the right-hand side of eqs.~(\ref{eq:mass}) --
(\ref{eq:total_energy}), while the transport terms are on the
left-hand side of these equations and are responsible for the
advection of mass, momentum and energy across the grid.

The \zeus\ method uses a von Neumann-Richtmyer artificial viscosity to
smooth shock discontinuities that may appear in fluid flows and can
cause a breakdown of the finite difference equations.  The artificial
viscosity term is added in the source terms as:
\begin{eqnarray}
\rho \frac{\partial\vecv}{\partial t} &=& - \grad p - \rho \grad \phi 
- \div \textbf{Q} \\
\frac{\partial e}{\partial t} &=& -p \div \vecv - \textbf{Q} : \grad \vecv, 
\end{eqnarray}
Here \textbf{Q} is the artificial viscosity stress tensor, which we
take to be diagonal with on-axis terms given by $l^2 \rho (\partial v
/ \partial x)^2$ as proposed by von Neumann \& Richtmyer.  The length
scale $l$ determines the width of shocks and is typically a few times
the cell spacing.

In our Cartesian coordinate system, the finite difference version of
these equations is particularly simple, although there is one
important complication.  In the \zeus\ formalism, the velocity is a
face-centered quantity -- that is, the velocity is recorded on a grid
that is staggered as compared to the density, pressure and energy,
which are at the cell center.  Therefore we must remember that $v_j$
is at position $x_{j-1/2}$ (we use this notation, rather than writing
$v_{j+1/2}$ both to match the original \zeus\ paper and also to make
it easier to compare these equations to what is actually in the code).

As in the original \zeus\ paper, the source terms are added in three
steps. First we add the pressure and gravity forces:
\begin{equation}
v_j^{n+a}  =  v_j^n - \frac{\Delta t}{\Delta x_j} \frac{p^n_j - p^n_{j-1}} {(\rho^n_j + \rho^n_{j-1})/2}.
\end{equation}
Partial updates are denoted by the $n+a$, $n+b$ notation.  We show
updates in one dimension as the extension to the multi-dimensional
case is straightforward (note that all dimensions are carried out for
each substep before progressing to the next substep). We then add the
artificial viscosity:
\begin{eqnarray}
v_j^{n+b} & = & v_j^{n+a} - \frac{\Delta t}{\Delta x_j} 
                             \frac{q^{n+a}_j - q^{n+a}_{j-1}} {(\rho^n_j + \rho^n_{j-1})/2} \\
e_j^{n+b} & = & e_j^n - \frac{\Delta t}{\Delta x_j} q^{n+a}_j (v^{n+a}_{j+1} - v^{n+a}_{j}).
\end{eqnarray}
The artificial viscosity coefficient $q_j$ is given by:
\begin{equation}
q_j = \left\{ \begin{array}{ll}
              Q_{\rm AV} \rho_j (v_{j+1} - v_j)^2 \quad & \rm{if}~(v_{j+1} - v_j) < 0 \\
               0 & \rm{otherwise}
               \end{array} \right.
\end{equation}
where $Q_{\rm AV}$ is a constant with a typical value of 2. We refer
the interested reader to \citet{Stone92a} and
\citet{1994ApJ...429..434A} for more details.  We also include the
option (turned off by default) of adding a linear artificial viscosity
as suggested in the \zeus\ paper for stagnant flow regions.  This is
given by
\begin{equation}
q_{{\rm lin},j} = Q_{\rm LIN} \rho c_j (v_{j+1} - v_{j})
\end{equation}
where $c_j^2 = \gamma p/\rho$ is the adiabatic sounds speed.

Finally, the third source step is the compression term and is given by
\begin{equation}
e^{n+c}_j = e^{n+b}_j \left( \frac{1 - (\Delta t/2) (\gamma - 1) (\div \vecv)_j }
                           {1 + (\Delta t/2) (\gamma - 1) (\div \vecv)_j } \right)
\end{equation}
We have used the notation $(\div \vecv)_j$ to indicate the
(potentially) multi-dimensional velocity divergence evaluated at the
cell center position $x_j$.  This equation differs from the previous
ones in that in the multi-dimensional case, still only one finite
difference equation is evaluated, but the divergence becomes
multi-dimensional.

We next examine the transport step, which is conservative.  Once
again, we dimensionally split the equations and present only the
one-dimensional version.  The finite difference equations actually
solved are:

\begin{equation}
\rho_j^{n+d} = \rho_j^{n} - \frac{\Delta t}{\Delta x} (v^{n+c}_{j+1/2} \rho^{*}_{j+1/2} - v^{n+c}_{j-1/2} \rho^{*}_{j-1/2} )
\end{equation}

Here $\rho^*_j$ is the correctly upwinded value of $\rho$ evaluated at
the cell-face corresponding to $v_j$, making $\rho^*_j v_j$ the mass
flux at the cell boundary and guaranteeing mass conservation.   This
requires interpolating each cell-centered quantity to the cell edge.
As recommended in \citet{Stone92a}, we use the second-order van Leer
scheme, which uses piecewise linear functions.  These are given by
Equations (48) and (49) of \citet{Stone92a}.  The transport steps for
the other variables are similar.  Note that we advect the specific
energy and specific momenta using the mass flux, as dictated by the
principle of consistent transport.  This requires appropriate
averaging for the momenta in the perpendicular directions as outlined
in equations (57)-(72) of \citet{Stone92a}.

A limitation of a technique that uses an artificial viscosity is that,
while the correct Rankine-Hugoniot jump conditions are achieved,
shocks are broadened over 3-4 mesh cells. This may cause unphysical
pre-heating of gas upstream of the shock wave, as discussed in
\citet{1994ApJ...429..434A}.  On the other hand, it is much more
robust than PPM and is easy to add additional physics.  We also note
that this method solves only the internal energy equation rather than
total energy, so the dual energy formulation discussed in
Section~\ref{sec.hydro.ppm} is unnecessary.